---
title: "DISH yield results"
subtitle: "preliminary"
author: "Sustainable Cropping Systems Lab UMN"
date: "`r Sys.Date()`"
format: pdf
editor: visual
execute:
  echo: false
  message: false
  warning: false
editor_options: 
  chunk_output_type: console
---

```{r conversions}
library(tidyverse)
library(measurements)

```

```{r}
read.csv("DISH master - yield (9).csv") -> dat

```

# Grain yield

```{r}
dat %>% 
  ggplot(aes(grain_combine_yield_uncleaned_kgha)) + 
  stat_bin()

# dat %>% 
#   filter(grain_combine_yield_uncleaned_kgha<100)
# this plot was noted as a poor stand

```

-   The one datapoint below 100 kg ha was noted during combine harvest as having a exceptionally poor stand of kernza. This plot received a 2X rate of prowl.

```{r}
dat %>% 
  ggplot(aes(treatment,grain_combine_yield_uncleaned_kgha)) +
  stat_summary() +
  geom_hline(yintercept = 360.93,linetype=3) +
  theme(axis.text.x = element_text(angle = 90)) +
  labs(xlab="", caption = "dotted line is control")

```

\newpage

# Forage yield

```{r}
dat %>% 
  ggplot(aes(forage_dmyield_Mg_ha)) + 
  stat_bin() +
  facet_wrap(~timing)
```

-   fall sprayed harvested in early May, spring sprayed harvested in late May.

\newpage

```{r}

# dat %>% 
#   filter(treatment=="control_0") %>% 
#   summarise(m=mean(grain_combine_yield_uncleaned_kgha))

dat %>% 
  filter(timing=="fall") %>% 
  ggplot(aes(treatment,forage_dmyield_Mg_ha, col=timing)) +
  stat_summary() +
  geom_hline(yintercept = 0.547, linetype = 3) +
  theme(axis.text.x = element_text(angle = 90)) +
  labs(x="")

dat %>% 
  filter(timing=="spring") %>% 
  ggplot(aes(treatment,forage_dmyield_Mg_ha, col=timing)) +
  stat_summary() +
  geom_hline(yintercept = 2.67, linetype = 3) +
  scale_colour_brewer(type = "div") +
  theme(axis.text.x = element_text(angle = 90)) +
  labs(x="")
```

-   positive effect from increasing rates of outrider for fall

-   negative effect of increasing finesse in fall

-   positive effect increasing axial in spring
