---
title: "DISH update"
author: "Jesse Puka-Beals"
date: "`r Sys.Date()`"
format: pdf
editor: visual
execute:
  echo: false
  message: false
  warning: false
editor_options: 
  chunk_output_type: console
bibliography: dish.bib
---

# Summary

Forage yields for fall sprayed were much lower than anticipated. Plots were harvested at a typical time (6May) and nodes had been observed a couple days prior in nearby fields. Once nodes are above 3 inches it is too late. This harvest can be considered at a typical time of year and maybe a little bit early for node height. 

We expected a forage yield of 2.4 Mg ha and observed 0.6 Mg ha. 

Data was double checked for accuracy. 

Water content of forage yield was 40%, which is typical. 

No obvious plausible treatment differences were observed among forage yield observations, unless there is a mechanism for axial and dual improving forage yield through weed control that is also observed in weed control ratings. 

For spring sprayed plots, we expect much higher yields. 

# Key information

DISH = Debalin Isisdor Herbicide Trial. Trial tests kernza response to certain herbicides

[DISH drive folder](https://drive.google.com/drive/folders/1o-Uxzj_jdKeDWl_txSn17louCpe-P9ZO)

[DISH master with data and plot map](https://docs.google.com/spreadsheets/d/1HtKeSv93GQMWm8yx5QJRuqrUkAzYI577hPCDMTsKLhU/edit#gid=1999649668)

[DISH data processing files](https://github.com/jpukabeals/scs-lab_misc)

```{r, warning=F, message=F, echo=F}
source("dish_yield.R")
```

## Forage yield

We expect spring forage yields around 2.4 Mg Ha [@Hunter2020a]

```{r}
dat5 %>% 
  select(plot, treatment, dmyield_Mg_ha) %>% 
  summary()
```

we observe 0.6 Mg Ha average spring forage yield.

```{r}
dat5 %>% 
  ggplot(aes(dmyield_Mg_ha)) +
  stat_bin() +
  geom_vline(xintercept = 2.4,linetype=2) +
  labs(caption = "Spring forage yield is typically 2.4 Mg ha (Hunter et all 2020)")

dat5 %>% 
  ggplot(aes(dmyield_Mg_ha)) +
  geom_boxplot() +
  labs(caption = "outliers were double checked and are correct")
```

\newpage
## water content
```{r}
dat %>% 
  filter(timing=="fall") %>% 
  # mutate(water_content = (forage_quadrat_wet_grams_with_bag - forage_quadrat_dry_grams_with_bag) / (forage_quadrat_dry_grams_with_bag + (forage_quadrat_wet_grams_with_bag - forage_quadrat_dry_grams_with_bag)) * 100)
  mutate(water_content = (forage_quadrat_wet_grams_with_bag - forage_quadrat_dry_grams_with_bag) /   forage_quadrat_wet_grams_with_bag * 100) %>% 
  ggplot(aes(water_content)) + 
  geom_boxplot() +
  labs(caption = "the outlier had a very small amount of sample but the value is correct")

```

water content averaged around 40%

## treatment differences

```{r}
dat5 %>% 
  # glimpse()
  ggplot(aes(herbicide,dmyield_Mg_ha)) +
  stat_summary() +
  geom_hline(yintercept=0.547, linetype=2) +
  labs(caption = "finesse and outrider may have lower DM yield than control")

dat5 %>% 
  # glimpse()
  mutate(rate = as.factor(rate)) %>% 
  ggplot(aes(rate,dmyield_Mg_ha)) +
  stat_summary() +
  labs(caption = "2x rates have similar yield as control, likely no herbicide injury from fall herbicide applicaiton on spring forage yield",
       x = "herbicide rate. control is 0, 2 is a 2X rate")

```


# References
