---
title: "SSD Spring Seeding Date Trial"
author: "Jesse Puka-Beals"
date: "`r Sys.Date()`"
format: pdf
editor: visual
execute:
  echo: false
  message: false
  warning: false
editor_options: 
  chunk_output_type: console
---

```{r}
source("ssd.R")
```

This trial occurred in K8 in the spring of 2024 to determine kernza establishment in early spring and whether different soil conditions resulted in different stand counts

Stand counts were a 12" length set randomly in a row and moved to a more representative area only if there was an obvious error in the section of row it was placed

Plots in the 500s did the best, with 508 and 509 as favorites. These are plots in uncompacted soil that were rototilled, packed, seeded, and then packed again with direction of packing not mattering. 

Takeaway: Soil prep maybe impacts seed germination but it wasn't observed in this trial. 

# Treatment list
```{r}
trt_codes %>% 
  knitr::kable()
```
 \newpage
```{r}
dat2 %>%
  group_by(treatment) %>% 
  tally() %>% 
  knitr::kable(caption = "there were 3 to 9 observations of every treatment")
```

# plant counts
```{r}
dat2 %>% 
  ggplot(aes(plants.per.ft)) +
  # stat_bin()
  geom_boxplot() +
  labs(caption = "need to remove 2 outliers")

dat2 %>% 
  filter(plants.per.ft < 28) -> dat3 

```

\newpage 
```{r}
dat3 %>%
  group_by(treatment) %>% 
  tally() %>% 
  knitr::kable(caption = "there were 3 to 9 observations of every treatment even after removing outliers")
```


# summary 

```{r}
dat3 %>% 
  group_by(treatment) %>% 
  summarise(mean = mean(plants.per.ft),
            sd = sd(plants.per.ft)) %>% 
  arrange(mean) %>% 
  mutate(mean = round(mean),
         sd = round(sd)) %>% 
  knitr::kable(caption = "Plants per foot summarized among different treatments")
```

```{r}
dat3 %>% 
  mutate(treatment = fct_reorder(factor(treatment),plants.per.ft)) %>% 
  ggplot(aes(treatment,plants.per.ft)) +
  geom_point(alpha=.5, shape=15) +
  stat_summary() +
  labs(caption = "Dots and bars are mean and standard error. squares are observations")
```

```{r}
dat3 %>% 
  lm(plants.per.ft~factor(treatment),.) %>% 
  anova() %>% 
  knitr::kable()
```


