---
title: "Short corn"
subtitle: "update"
author: "Jesse Bealsburg"
date: "`r Sys.Date()`"
format: pdf
editor: visual
execute:
  echo: false
  message: false
  warning: false
editor_options: 
  chunk_output_type: console
---

### Notes

DKC is the check variety that is not a short stature corn hybrid.

Moisture content and test weight determined using a Perten 5200

\newpage

# Growth monitoring

### Figures

```{r}
library(tidyverse)
read.csv("Cor Datasheet - data (9).csv") -> dat

dat %>% 
  # select(-stage) %>% 
  rename(plant_height = plant.height..cm.,
         leaf_count = leaf.count,
         stem_circumference = stem.circumference..cm.,
         light = light.penetration,
         node_number = node.number,
         node_tassel_distance = distance.between.node.and.tassel..cm.) %>% 
  mutate(plot = factor(plot),
         date = as.POSIXct(date,
                           format = "%m/%d/%Y"))-> dat
```

```{r}
dat %>% 
  ggplot(aes(date,plant_height,
             col=name)) +
  stat_summary(geom = "line") +
  stat_summary() +
  labs(y="plant height (cm)")
```

-   Person collecting data on 6Aug and 14Aug likely misinterpreted protocol for measuring plant heights

\newpage

```{r}
dat %>% 
  ggplot(aes(date,leaf_count,
             col=name,
             shape = name)) +
  stat_summary(geom = "line",
               linetype = 2) +
  stat_summary()
```

-   red dots (DKC56-65Rib) in august are right behind the blue and are therefore hidden/hard to see

\newpage

```{r}
dat %>% 
  ggplot(aes(date,stem_circumference,
             col=name)) +
  stat_summary(geom = "line") +
  stat_summary()
```

-   

\newpage

```{r}
dat %>% 
  ggplot(aes(date,node_number,
             col=name)) +
  stat_summary(geom = "line") +
  stat_summary()
```

-   

\newpage

```{r, include=F}

dat %>% 
  ggplot(aes(date,node_tassel_distance,
             col=name)) +
  stat_summary(geom = "line") +
  stat_summary() +
  labs(y="distance from 5th node to base of tassel (cm)")
```

-   

\newpage

```{r}
dat %>% 
  select(date,plot, name,stage) %>% 
  filter(stage != "") %>% 
  mutate(stagenum = str_sub(stage,2)) %>% 
  ggplot(aes(date,stagenum,
             col=name,
             group = name)) +
  stat_summary(geom = "line") +
  stat_summary(position = position_dodge(.4)) +
  labs(y="Reproductive R stage\n (i.e. R1 vs R2 vs R3)",
       x="") 
  ## need to work on this graph, need better seperation

dat %>% 
  select(date,plot, name,stage) %>% 
  filter(stage != "") %>% 
  mutate(stagenum = str_sub(stage,2)) %>% 
  ggplot(aes(date,stagenum,
             col=name)) +
  stat_summary(geom="pointrange",position = position_dodge(.4)) +
  labs(y="Reproductive R stage\n (i.e. R1 vs R2 vs R3)",
       x="") +
  facet_wrap(~date)
```

\newpage

# Harvest

## Yield

```{r}

# grain is from 2 meters at 30" row spacing
# conv_unit(30,"inch","m")
# 2*.762 is square meters, 1.524 square meters

# test weight is 56 lbs per bushel
library(measurements)
dat %>% 
  filter(collection=="harvest") %>% 
  select(date,plot,type,name,population,grain_dry_g,rowlength_totalcobcount) %>% 
  mutate(grain_extrap = grain_dry_g/5*rowlength_totalcobcount,
    yield_kgha = conv_multiunit(grain_extrap/1.524,
                                from = "g / m2",
                                to = "kg / hectare"),
         yield_buacre = conv_multiunit(yield_kgha,
                                       from = "kg / hectare",
                                       to = "lbs / acre")/56) -> dat2

dat2 %>% 
  ggplot(aes(yield_buacre)) + 
  geom_boxplot()

dat2 %>% 
  ggplot(aes(type,yield_buacre)) + 
  geom_boxplot()

dat2 %>% 
  ggplot(aes(type,yield_buacre,
             fill = name)) + 
  geom_boxplot() +
  labs(x="",
       y= "grain yield \n(bushels / acre)") +
  theme(legend.title = element_blank())

dat2 %>% 
  ggplot(aes(type,yield_kgha,
             fill = name)) + 
  geom_boxplot() +
  labs(x="",
       y= "grain yield \n(kilograms / hectare)") +
  theme(legend.title = element_blank())



dat2 %>% 
  summarise(n=n(),
            mean = mean(yield_kgha),
            sd = sd(yield_kgha),
            cv = sd/mean*100,
            max = max(yield_kgha),
            min = min(yield_kgha)) %>% 
  mutate(across(.cols = 2:6,round)) %>% 
  knitr::kable(caption = "predicted grain yield in kg ha")

dat2 %>% 
  summarise(n=n(),
            mean = mean(yield_buacre),
            sd = sd(yield_buacre),
            cv = sd/mean*100,
            max = max(yield_buacre),
            min = min(yield_buacre)) %>% 
  mutate(across(.cols = 2:6,round)) %>% 
  knitr::kable(caption = "predicted grain yield in bushels per acre at a test weight of 56 lbs per bushel which was determined at harvest")
```

\newpage

## Moisture content and test weight

```{r moisture content}
dat %>% 
  filter(collection=="harvest") %>% 
  select(#type,name,population,
         plot,
         grain_moisturePercent_afterdryer,
         grain_moisturePercent_atharvest,
         grain_moisturePercent_afterdryerandaspiration,
         ) %>% 
  pivot_longer(
    cols = c(grain_moisturePercent_atharvest,grain_moisturePercent_afterdryer,
             grain_moisturePercent_afterdryerandaspiration)
  ) %>% 
  mutate(timing = str_sub(name,start=23),
         data = str_sub(name, 7,15)) -> dat3

dat3 %>% 
  select(-name) %>% 
  left_join(dat2) %>%
  mutate(timing = factor(timing,levels = c("atharvest", "afterdryer", "afterdryerandaspiration" ))) %>% 
  ggplot(aes(timing,value,fill=plot,
             color = plot, 
             group = plot)) +
  geom_point() +
  geom_line() +
  labs(y="Moisture Content (%)",
       x="")

```

-   moisture content decreased in dryer to below 10%. After it's below 10% the machine prediction is less accurate

\newpage

```{r}
dat3 %>% 
  select(-name) %>% 
  left_join(dat2) %>%
  mutate(timing = factor(timing,levels = c("atharvest", "afterdryer", "afterdryerandaspiration" ))) %>% 
  ggplot(aes(timing,value,fill=name,
             color = name, 
             group = name)) +
  stat_summary() +
  stat_summary(geom = "line") +
  labs(y="Moisture Content (%)",
       x="")
```

\newpage

```{r}
dat %>% 
  filter(collection=="harvest") %>% 
  select(#type,name,population,
         plot,
         grain_testweight_lbbu_afterdryer,
         grain_testweight_lbbu_atharvest,
         grain_testweight_lbbu_afterdryerandaspiration
         ) %>% 
  pivot_longer(
    cols = c(grain_testweight_lbbu_afterdryer,
             grain_testweight_lbbu_atharvest,
            grain_testweight_lbbu_afterdryerandaspiration)
  ) %>% 
  mutate(timing = str_sub(name,start=23),
         data = str_sub(name, 7,15)) -> dat4

dat4 %>% 
  bind_rows(dat3) %>% 
  select(-name) %>% 
  left_join(dat2) %>% 
  mutate(timing = factor(timing,levels = c("atharvest", "afterdryer", "afterdryerandaspiration" ))) -> dat5

dat5 %>%
  filter(data == "testweigh") %>% 
  ggplot(aes(timing,value,fill=plot,
             color = plot, 
             group = plot)) +
  geom_point() +
  geom_line() +
  geom_hline(yintercept = 56, linetype=2) +
  labs(y="Test Weight (lb / bu)",
       x="")

```

-   All plots appear to have test weight increase after drying and then again after aspiration. This could be from gaining moisture from sitting longer, but I would think that would decrease test weight?

-   Plot 1 is weird, test weight does not change

\newpage

```{r}
dat5 %>%
  filter(data == "testweigh") %>% 
  ggplot(aes(timing,value,fill=name,
             color = name, 
             group = name)) +
  stat_summary() +
  stat_summary(geom = "line") +
  labs(y="Test Weight (lb / bu)",
       x="")
```
