---
title: "weather import"
author: "Jesse Puka-Beals"
date: '`r format(Sys.Date(), "%m %d %Y")`'
output: html_document
---

```{r install packages, include=FALSE, cache=TRUE}
knitr::opts_chunk$set(echo = TRUE)

if (!require("pacman")) install.packages("pacman")
pacman::p_load(tidyverse, dplyr)

```

```{r}
dat <- read.csv("~/R data/weather.nasa.csv", skip=12) #skip first 12 lines which explain the column headers

dat1 <- dat %>%
  mutate(year=YEAR,
         precip.mm=PRECTOT,
         tmax.c=T2M_MAX,
         tmin.c=T2M_MIN) %>%
  select(year,DOY,precip.mm,tmax.c,tmin.c) %>%
  mutate(tmean=(tmax.c+tmin.c)/2) 

```

now we need to mutate new columns to add GDD

Then we need to create a rule where the data is grouped within year starting at DOY=1
for each DOY, we need a cum.sum for precip.mm
for each DOY after the mean temp has been higher than 0 for 5 consecutive days, we need a cum.sum, restarting with the next year


```{r GDD}

dat2 <- dat1 %>%
  mutate(GDD=(tmax.c+tmin.c)/2-0)
head(dat2)

# add if, then statement where if GDD<=0, =0

```

> x[x<0] <- 5; x   # modify elements less than 0

> x
[1] -3 -2 -1  0  1  2
> x[2] <- 0; x        # modify 2nd element
[1] -3  0 -1  0  1  2
> x[x<0] <- 5; x   # modify elements less than 0
[1] 5 0 5 0 1 2
> x <- x[1:4]; x      # truncate x to first 4 elements
[1] 5 0 5 0

```{r}
x <- dat2$GDD 
x[x<0] <- 0
dat2$GDD <- x
head(dat2)
```



```{r}
dat2[7]
dat2$GDD

dat2[7<0]

dat2[7<0] <- 0

dat2$GDD[1<0] <- 0

dat2
```


if (test_expression) {
statement
}

```{r}
if (dat2$GDD<0) {
dat2$GDD=0
}

dat2$GDD
```
if (test_expression) {
statement1
} else {
statement2
}

```{r}
if (dat2$GDD<0) {
dat2$GDD=0
} else {
dat2$GDD=dat2$GDD
}
```
if ( test_expression1) {
statement1
} else if ( test_expression2) {
statement2
} else if ( test_expression3) {
statement3
} else {
statement4
}


