---
title: "Buckwheat variety trial 2024 MN"
subtitle: ""
author: "Jesse Bealsburg"
date: "`r Sys.Date()`"
format: pdf
editor: visual
execute:
  echo: false
  message: false
  warning: false
editor_options: 
  chunk_output_type: console
---

```{r import and clean, include=F}

read.csv("MASTER organic buckwheat 2024 - data_firstFlowers.csv") -> dat_flower
read.csv("MASTER organic buckwheat 2024 - data_greenSeeds.csv") -> dat_greenseed
read.csv("MASTER organic buckwheat 2024 - data_ripeSeeds.csv") -> dat_ripeseed

library(tidyverse)


bind_rows(dat_flower,dat_greenseed,dat_ripeseed) -> dat


dat %>% 
  rename(canopeo = canopeo....) -> dat

# Conversions
library(measurements)
# MN row spacing is 6 inch, 1 meter length. area is 0.1524 square meter 
conv_unit(6,"in","m")


dat %>% 
  mutate(seedyield_kgha = conv_multiunit(yield_dry_grain/.1524,
                                         from = "g / m2",
                                         to = "kg / hectare"),
         strawDMyield_Mgha = conv_multiunit(yield_dry_straw/.1524,
                                         from = "g / m2",
                                         to = "Mg / hectare"),
         watercontent_percent = (yield_wet_total-yield_dry_total)/yield_wet_total*100) -> dat
  

```

# Yield

takeaway: Koto yielded less than other varieties

```{r}
dat %>% 
  # distinct(timing)
  filter(timing == "ripe seeds") %>% 
  ggplot(aes(seedyield_kgha)) +
  stat_bin(bins = 7) +
  labs(caption = "1 missing data point")

dat %>% 
  # distinct(timing)
  filter(timing == "ripe seeds") %>% 
  ggplot(aes(variety,seedyield_kgha)) +
  stat_summary()

dat %>% 
  filter(timing == "ripe seeds") %>% 
  drop_na(seedyield_kgha) %>% 
  summarise(mean = mean(seedyield_kgha),
            max = max(seedyield_kgha),
            min = min(seedyield_kgha),
            n = n()) %>% 
  knitr::kable(caption = "seed yield in kg ha")

dat %>% 
  filter(timing == "ripe seeds") %>% 
  drop_na(seedyield_kgha) %>% 
  group_by(variety) %>% 
  summarise(mean = mean(seedyield_kgha),
            max = max(seedyield_kgha),
            min = min(seedyield_kgha),
            n = n()) %>% 
  knitr::kable(caption = "seed yield in kg ha")


```

-   Seed yield averaged 6921 kg ha

-   koto yielded less than other varieties

```{r}
dat %>% 
  # distinct(timing)
  filter(timing == "ripe seeds") %>% 
  ggplot(aes(variety,strawDMyield_Mgha)) +
  stat_summary()

dat %>% 
  filter(timing == "ripe seeds") %>% 
  drop_na(strawDMyield_Mgha) %>% 
  summarise(mean = mean(strawDMyield_Mgha),
            max = max(strawDMyield_Mgha),
            min = min(strawDMyield_Mgha),
            n = n()) %>% 
  knitr::kable(caption = "straw dry matter yield in Mg ha")
```

-   straw dry matter yield averaged around 17 Mg ha

-   straw dry matter yield was variable with no obvious differences among varieties

# Weed suppression

Takeaway: koto was shorter and had more leaves but had similar soil coverage as other varieties on 29July.

```{r}
dat %>% 
  filter(timing == "first flowering") %>%    
  ggplot(aes(variety,canopeo)) +   
  stat_summary()  
```

-   On 23July, devyatka seems to be closing canopy slower than other buckwheat varieties

```{r}
dat %>% 
  filter(timing == "first flowering") %>%    
  ggplot(aes(variety,height)) +   
  stat_summary() +
  labs(y="plant height\n(cm)")
```

-   On 23July, Koto is shorter than other buckwheat varieties

```{r}
dat %>% 
  filter(timing == "first flowering") %>%    
  ggplot(aes(variety,leaf_count)) +   
  stat_summary() 
```

-   On 23July, Koto may average more leaves than other varieties per plant.

```{r}
dat %>% 
  # distinct(timing)
  filter(timing == "ripe seeds") %>% 
  ggplot(aes(variety,weed_score)) +
  stat_summary() +
  labs(caption = "collected at harvest on 16Sep20204. Higher score means more weedy.")

```

-   At harvest, weediness is an indirect way of measuring how patchy the stand was. Patchiness coming from lodging. Koto seems to be weedier than at least glgf land race, which would also agree with koto having a lower seed yield. Looks like koto stands just didn't do that good.
